---
title: "esm206-lab6-cje"
author: "Chance English"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


library(tidyverse)
library(ggplot2)
library(palmerpenguins)
library(broom)
library(equatiomatic)


```



## Example of a rank-based test 


We'll make out own samples using pseudo random number generator 


```{r}

set.seed(1414) # sets a starting point from which we ask it to return sample values. Allows comparrison among other people

gp1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)

gp2 <- sample.int(30, size = 15, replace = TRUE)

gp_df <- data_frame(gp1, gp2)

gp_df_tidy <- gather(gp_df, "gp", "random_number")
```




```{r}

t.test(gp1, gp2)

```

Conclusion: The probability of finding at least out different means for gp1 and gp2 if they were drawn from the same population is 19.8%. We fail to reject the null hypothesis. There is no significant difference between group 1 and group 2.

Now let's compare thus outcome to a rank based test. 


## Mann Whitney U unpaired rank-based test 

```{r}
 
mwu <- wilcox.test(gp1, gp2)

```

If these samples were drawn from populations with the same ranks (median) the probability of finding two samples with ranks *at least as different* as those in our samples is 0.28. We fail to reject the null hypothesis. There is no significant difference in ranks between group 1 and group 2. Median scores for group 1 (M = 14) and group 2 (M = 2) did not differ significantly (Mann Whitney U test : U(df) = 86, p = 0.28)



## Linear Regression 

Simple linear regression: single dependent variable, a single dependent independent variable 


```{r, include = FALSE}

penguins %>% ggplot(aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_abline(slope = 49.686, intercept = -5780.831)


```

Find a linear regression model using ordinary least squares describing the relationship between flipper length and body mass for these penguins 


3 Pieces

- What type of model?
- What is the relationshio to model (DV - IV's)
- Where is the data that is used to create this model?


```{r}

penguin_lm <- lm(data = penguins, body_mass_g ~ flipper_length_mm)



```

## Intercept would be a negative mass. All models are wrong, and extrapolation can be difficult



### broom package retunrs model outputs as tidy dataframes 


```{r}

penguin_lm_tidy <- broom::tidy(penguin_lm)


```

How can I include my model equation in a report?



```{r}
extract_eq(model = penguin_lm, use_coef = TRUE)


```



```{r}

plot(penguin_lm)

```

