---
title: "esm206-lab6-cje"
author: "Chance English"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(ggplot2)
library(palmerpenguins)
library(broom)
library(equatiomatic)


```



## Example of a rank-based test 


We'll make out own samples using pseudo random number generator 


```{r}

set.seed(1414) # sets a starting point from which we ask it to return sample values. Allows comparrison among other people

gp1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)

gp2 <- sample.int(30, size = 15, replace = TRUE)

gp_df <- data_frame(gp1, gp2)

gp_df_tidy <- gather(gp_df, "gp", "random_number")
```


```{r}

hist(gp

```


```{r}

t.test(gp1, gp2)

```

Conclusion: The probability of finding at least out different means for gp1 and gp2 if they were drawn from the same population is 19.8%. We fail to reject the null hypothesis. There is no significant difference between group 1 and group 2.

Now let's compare thus outcome to a rank based test. 


## Mann Whitney U unpaired rank-based test 

```{r}
 
mwu <- wilcox.test(gp1, gp2) # it we have ties, it will make an approximation in order to compute ranks. 


mwu
```

If these samples were drawn from populations with the same ranks (median) the probability of finding two samples with ranks *at least as different* as those in our samples is 0.28. We fail to reject the null hypothesis. There is no significant difference in ranks between group 1 and group 2. Median scores for group 1 (M = 14) and group 2 (M = 2) did not differ significantly (Mann Whitney U test : U(df) = 86, p = 0.28)
